# (APPENDIX) Appendix {#Appendix}

## Appendix A: Open-source Software

### Open-Source Software

#### QGIS

#### R and RStudio

#### RS Minerve

## Appendix B: Quick Guides

## Appendix C: Processing Climate Data

### Cutting CHELSA v.1.2.1 to CA Domain

The CHELSA v1.2.1 data is a global dataset. The data can be cut to the Central Asia domain in the following way.

```{r,eval=FALSE}
dir_CHELSA <- '/Users/tobiassiegfried/Dropbox (hydrosolutions)/1_HSOL_PROJECTS/PROJECTS/SDC/DKU_WRM_COURSE_CA/CourseMaterials/Data/CLIMATE/CHELSA_V1_2_1/'

# First, we process precipitation data (prec).
prec_CHELSA_files <- list.files(paste0(dir_CHELSA,'tp/'))
for (idx in 1:length(prec_CHELSA_files)){
  str2Print <- paste0('Processing PREC: ', idx, ' out of ', length(prec_CHELSA_files) )
  print(str2Print)
  fName <- prec_CHELSA_files[idx]
  global_CHELSA_raster <- raster(paste0(dir_CHELSA,'prec/',fName))
  centralAsia_CHELSA_raster <- raster::crop(global_CHELSA_raster,aoi_CentralAsia_LatLon)
  # Save the centralAsia_CHELSA_raster
  fName2Save <- paste0(substr(fName,1,19),'_CA',substr(fName,20,30))
  raster::writeRaster(centralAsia_CHELSA_raster,
                      paste0(dir_CHELSA,'prec/',fName2Save),
                      'GTiff',
                      overwrite = TRUE)
}

# Second, we process mean temperature data (tmean).
tmean_CHELSA_files <- list.files(paste0(dir_CHELSA,'t2m/'))
for (idx in 1:length(tmean_CHELSA_files)){
  str2Print <- paste0('Processing TMEAN: ', idx, ' out of ', length(tmean_CHELSA_files) )
  print(str2Print)
  fName <- tmean_CHELSA_files[idx]
  global_CHELSA_raster <- raster(paste0(dir_CHELSA,'tmean/',fName))
  centralAsia_CHELSA_raster <- raster::crop(global_CHELSA_raster,aoi_CentralAsia_LatLon)
  # Save the centralAsia_CHELSA_raster
  fName2Save <- paste0(substr(fName,1,20),'_CA',substr(fName,21,35))
  raster::writeRaster(centralAsia_CHELSA_raster,
                      paste0(dir_CHELSA,'tmean/',fName2Save),
                      'GTiff',
                      overwrite = TRUE)
}

(raster(paste0(dir_CHELSA,'tmean/',fName2Save))/10 - 273.15) %>% plot()
```

### Bias Correcting CHELSA v1.2.1 Precipitation Data for Snow Undercatch

```{r,eval=FALSE}
## general info - function arguments
basinName <- 'Gunt'
dataType_ERA5 <- 'tp'
#basinShape <- 'to be passed in'
targetCRS <- "+init=epsg:32642"
## Directories of relevant climate files  - function arguments
dir_CHELSA <- '/Users/tobiassiegfried/Dropbox (hydrosolutions)/1_HSOL_PROJECTS/PROJECTS/SDC/DKU_WRM_COURSE_CA/CourseMaterials/Data/CLIMATE/CHELSA_V1_2_1/'
# Basin AOI - function arguments, but check if all can be derived from the basinShape that is anyhow to be passed in!
aoi_Basin_LatLon <- gunt_Shapefile_LatLon %>% extent # GUNT
# ===
fileCorrFact <- '/Users/tobiassiegfried/Dropbox (hydrosolutions)/1_HSOL_PROJECTS/PROJECTS/SDC/DKU_WRM_COURSE_CA/CourseMaterials/Data/CLIMATE/PBCOR_V1/CHELSA_V12.nc'
# Get CHELSA file list - Note, we already have cut to the CA domain!
fileList <- list.files(paste0(dir_CHELSA,dataType_ERA5))
# beginning and end
startY <- 1981
endY <- 2013
# Monthly correction factors (Beck et al., 2020)
#pbcorr_monthly <- nc_open(dir_CorrFact)
pbcorr_monthly <- brick(fileCorrFact, varname="corr_fac_monthly")
pbcorr_monthly_basin_longlat <- raster::crop(pbcorr_monthly,aoi_CentralAsia_LatLon)
# start to loop through years and months for bias correcting the monthly values.
for (yr in startY:endY){
  for (mon in 1:12){
    # Load corresponding tmean CHELSA Central Asia File
    if (mon<10){
      chelsa_data_orig <- 
        raster(paste0(dir_CHELSA,dataType_ERA5,'/CHELSA_prec_',yr,'_0',mon,'_CA_V1.2.1.tif'))
      numbPrefix <- '_0'
    } else {
      chelsa_data_orig <- 
        raster(paste0(dir_CHELSA,dataType_ERA5,'/CHELSA_prec_',yr,'_',mon,'_CA_V1.2.1.tif'))
      numbPrefix <- '_'
    }
    chelsa_data_orig_resamp <- raster::resample(chelsa_data_orig,pbcorr_monthly_basin_longlat)
    chelsa_data_bcorr_resamp <- overlay(chelsa_data_orig_resamp,subset(pbcorr_monthly_basin_longlat,mon),
                                        fun=function(x,y){return(x*y)})
   # write bias corrected CHELSA back to the disk.
    raster::writeRaster(chelsa_data_bcorr_resamp,
                        paste0(dir_CHELSA,dataType_ERA5,'/CHELSA_tp_bcorr_',yr,numbPrefix,mon,'_CA_V1.2.1.tif'),
                        'GTiff',
                        varname   = dataType_ERA5,
                        overwrite = TRUE)
    #chelsa_data_bcorr_resamp %>% plot()
  }
}
```

### Computing Mean Monthly Temperature Climatology from CHELSA v1.2.1 data

The following code take CHELSA_V1.2.1 mean monthly temperature raster data and computes long-term mean monthly temperatures over the Central Asia domain. The resulting file can optionally be stored on the disk.

```{r, eval=FALSE}
fDir <- '../HydrologicalModeling_CentralAsia_Data/CentralAsiaDomain/CHELSA_V1.2.1/'
# First, we process and display mean monthly 2 meters temperatures.
varDir <- 't2m/'
t2m_files <- list.files(paste0(fDir,varDir))
# Get monthly fields and average over all data for resulting climatology
idxSeq <- seq(1,409,12)
for (idx in (1:12)){
  print(months[idx])
  t2m_files_sel <- t2m_files[idxSeq]
  for (idxRaster in (1:length(t2m_files_sel))){
    if (idxRaster == 1){
      raster_res <- raster::raster(paste0(fDir,varDir,'/',t2m_files_sel[1]))
    } else {
      raster_n <- raster::raster(paste0(fDir,varDir,'/',t2m_files_sel[idxRaster]))
      raster_res <- raster::addLayer(raster_n,raster_res)
    }
    raster_res_mean <- calc(raster_res, fun = mean)
  }
  if (idx ==1){
    raster_res_mean_stack <- raster_res_mean
  } else {
    raster_res_mean_stack <- raster::addLayer(raster_res_mean_stack,raster_res_mean)
  }
  idxSeq  <- idxSeq + 1
}

names(raster_res_mean_stack) <- month.abb

# raster::writeRaster(raster_res_mean_stack,
#                         paste0(fDir,'t2m_climatology/t2m_climatology_CA.tiff'),
#                         options="INTERLEAVE=BAND",
#                         'GTiff',
#                         varname   = 't2m',
#                         overwrite = TRUE)
```

### ERA5
