---
title: "_package_maintenance"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Package Data Maintenance

Install Package from GitHUB
```{r}
install_github("tobiassiegfried/riversCentralAsia")
```

## Data Preparation for Data Accessible by users

We make data available to the users in different basin-specific data files and call them accordingly

-   ChirchikRiverBasin
-   SyrDaryaBasin (to come)
-   AmuDaryaBasin (to come)
-   ChuRiverBasin (to come)
-   TalasRiverBasin (to come)

For this preparatory work here, we need the following libraries

```{r}
# Load Package
library(devtools)
library(tidyverse)
library(lubridate)    
library(timetk)       
library(riversCentralAsia)
```

### Chirchik River Basin
#### Discharge Data
```{r}
# Chirchik River
fPath <- '/Users/tobiassiegfried/Dropbox (hydrosolutions)/1_HSOL_PROJECTS/PROJECTS/SDC/DKU_WRM_COURSE_CA/Course Materials/Data/Q/ChirchikRiverBasin/'
## Chatkal River, Khudaydod
Q_16279 <- loadTabularData(fPath,'16279_decadal.csv',16279,'Khudaydod','Chatkal','Chirchik','Q','m3s')
## Pskem River, Mullala
Q_16290 <- loadTabularData(fPath,'16290_decadal.csv',16290,'Mullala', 'Pskem','Chirchik','Q','m3s')
## Inflow Charvak
Q_16924 <- loadTabularData(fPath,'16924_decadal.csv',16924,'Lake_Inflow','Chatkal','Chirchik','Q','m3s')
## Nauvalisoy River, Sidzhak
Q_16298 <- loadTabularData(fPath,'16298_decadal.csv',16298,'Sidzhak','Nauvalisoy','Chirchik','Q','m3s')
## Chirchik River, Chinaz
Q_16275 <- loadTabularData(fPath,'16275_monthly.csv',16275,'Chinaz','Chirchik','Chirchik','Q','m3s')
## Ugam River, Khodizhikent
Q_16300 <- loadTabularData(fPath,'16300_monthly.csv',16300,'Khodzhikent','Ugam','Chirchik','Q','m3s')
## Chirchik, Gazalkent
Q_16262 <- loadTabularData(fPath,'16262_monthly.csv',16262,'Gazalkent','Chirchik','Chirchik','Q','m3s')

## Assemble Chirchik River Basin discharge data
data.Q.ChirchikRiverBasin <- 
  bind_rows(Q_16279,Q_16290,Q_16924,Q_16298,Q_16300,Q_16275,Q_16262)
```

#### Meteostation Data
```{r}
fPath <- '/Users/tobiassiegfried/Dropbox (hydrosolutions)/1_HSOL_PROJECTS/PROJECTS/SDC/DKU_WRM_COURSE_CA/Course Materials/Data/METEO/ChirchikRiverBasin/'

## Pskem Meteo Station
P_38462 <- loadTabularData(fPath,'38462_P_decadal.csv',38462,'Pskem Meteo Station','Pskem','Chirchik','P','mm')
T_38462 <- loadTabularData(fPath,'38462_T_decadal.csv',38462,'Pskem Meteo Station','Pskem','Chirchik','T','degC')

## Charvak Reservoir Meteo Station
P_38464 <- loadTabularData(fPath,'38464_P_decadal.csv',38464,'Charvak Reservoir Meteo Station','Chirchik','Chirchik','P','mm')

## Chatkal Meteo Station
P_38471 <- loadTabularData(fPath,'38471_P_decadal.csv',38471,'Chatkal Meteo Station','Chatkal','Chirchik','P','mm')
T_38471 <- loadTabularData(fPath,'38471_T_decadal.csv',38471,'Chatkal Meteo Station','Chatkal','Chirchik','T','degC')

## Oygaing Meteo Station
P_38339 <- loadTabularData(fPath,'38339_P_mon.csv',38339,'Oygaing Meteo Station','Pskem',
                           'Chirchik','P','mm')

# Assemble the data
data.PT.ChirchikRiverBasin <- bind_rows(P_38462,T_38462,P_38464,P_38471,T_38471,P_38339)
```

Bind all data
```{r}
data.ChirchikRiverBasin <- bind_rows(data.Q.ChirchikRiverBasin,data.PT.ChirchikRiverBasin)
```

#### Chirchik River: Station information
```{r}
stationData <- data.ChirchikRiverBasin %>% select(code) %>% unique()

stationData$lon_UTM42 <- c(
                            598278, 
                            597351, 
                            584616, 
                            589674, 
                            578612, 
                            479463,
                            564916,
                            613221.09, 
                            586345.41, 
                            684252.58,
                            654406.23)

stationData$lat_UTM42 <- c(
                            4596628, 
                            4622724,
                            4609108, 
                            4618690, 
                            4610070, 
                            4528236,
                            4602785,
                            4641209.20, 
                            4607704.44,
                            4636113.34,
                            4669602.08)

stationData$altitude_masl <- c(
                            953, 
                            911, 
                            864, 
                            941, 
                            759, 
                            254,
                            681,
                            1136, 
                            970, 
                            1185,
                            2177)

stationData$basinSize_sqkm <- c(
                                  5677829056, 
                                  2521388472, 
                                  9999165900, 
                                  98555481, 
                                  864946131, 
                                  13112062045,
                                  11139775526,
                                  NA,
                                  NA,
                                  NA,
                                  NA)
stationData$basinSize_sqkm <- stationData$basinSize_sqkm / 10^6
stationData
```

Complete Chirchik River Basin
```{r}
data.ChirchikRiverBasin <- full_join(data.ChirchikRiverBasin,stationData,by='code')
```

#### Store Chirchik River Basin Data (see then _PackageDataMaintenance.RMD in Package Folder for package data deployment)
```{r}
ChirchikRiverBasin <- data.ChirchikRiverBasin
fLoc <- '/Users/tobiassiegfried/Dropbox (hydrosolutions)/1_HSOL_PROJECTS/PROJECTS/SDC/DKU_WRM_COURSE_CA/Course Materials/Handbook/Applied_Hydrological_Modeling_Bookdown/data/ChirchikRiverBasin.rda'
saveRDS(ChirchikRiverBasin,fLoc)
```

### Syr Darya Basin

### Amu Darya Basin

### Chu River Basin

### Talas River Basin

