---
title: "_scratch_ClimateChange"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load Required Libraries
```{r,echo=FALSE}
library(devtools)
library(tidyverse)
library(lubridate)    
library(timetk)       
library(riversCentralAsia)
```


# Investigating a Changing Climate in Central Asia
Load data
```{r}
fPath <- '/Users/tobiassiegfried/Dropbox (hydrosolutions)/1_HSOL_PROJECTS/PROJECTS/SDC/DKU_WRM_COURSE_CA/Course Materials/Handbook/Applied_Hydrological_Modeling_Bookdown/temp/'
data_wide_tbl <- readRDS(paste(fPath,'data_wide_tbl',sep=""))
data_long_tbl <- readRDS(paste(fPath,'data_long_tbl',sep=""))
```

## Chirchik River

### Pskem 
#### 16290 - Q
##### Monthly
```{r,message=FALSE,warning=FALSE}
pl <- data_long_tbl %>% filter(name=='Q16290') %>% 
  summarise_by_time(.date_var = date, .by="month",value=mean(value)) %>% 
  tk_ts(frequency = 12) %>% 
  forecast::ggsubseriesplot(year.labels = FALSE) + 
              geom_smooth(method = "lm",color="red") +
              ggtitle('Mean mon. discharge (Station 16290, 1933 - 2015)') +
              xlab('month') +
              ylab('m^3/s')

fPath <-  '/Users/tobiassiegfried/Dropbox (hydrosolutions)/1_HSOL_PROJECTS/PROJECTS/SDC/DKU_WRM_COURSE_CA/Course Materials/Data/Q/ChirchikRiverBasin/'
ggsave(paste(fPath,'16290_Pskem_Q_SubSeriesPlot.pdf',sep=""),
       device ="pdf",
       width  = 6,
       height = 2.5)
```
##### Quarter
```{r,message=FALSE,warning=FALSE}
pl <- data_long_tbl %>% filter(name=='Q16290') %>% 
  summarise_by_time(.date_var = date, .by="quarter",value=mean(value)) %>% 
  tk_ts(frequency = 4) %>% 
  forecast::ggsubseriesplot(year.labels = FALSE) + 
              geom_smooth(method = "lm",color="red") +
              ggtitle('Mean mon. discharge (Station 16290, 1933 - 2015)') +
              xlab('month') +
              ylab('m^3/s')

pl

# 
# fPath <-  '/Users/tobiassiegfried/Dropbox (hydrosolutions)/1_HSOL_PROJECTS/PROJECTS/SDC/DKU_WRM_COURSE_CA/Course Materials/Data/Q/ChirchikRiverBasin/'
# ggsave(paste(fPath,'16290_Pskem_Q_SubSeriesPlot.pdf',sep=""),
#        device ="pdf",
#        width  = 6,
#        height = 2.5)
```

Get long term average Pskem River discharge
```{r}
data_long_tbl %>% filter(name=='Q16290') %>% summarise(Average=mean(value,na.rm=TRUE))
```

#### 38462 - T
T: Pskem Meteorological Station (38462)
##### Monthly T
```{r,message=FALSE,warning=FALSE}
pl <- data_long_tbl %>% filter(name=='T38462') %>% 
  summarise_by_time(.date_var = date, .by="month",value=mean(value)) %>% 
  tk_ts(frequency = 12) %>% 
  forecast::ggsubseriesplot(year.labels = FALSE) + 
              geom_smooth(method = "lm",color="red") +
              ggtitle('Mon. mean temperature (Station 38462, 1933 - 2015)') +
              xlab('month') +
              ylab('deg. C.')

fPath <-  '/Users/tobiassiegfried/Dropbox (hydrosolutions)/1_HSOL_PROJECTS/PROJECTS/SDC/DKU_WRM_COURSE_CA/Course Materials/Data/METEO/ChirchikRiverBasin/'
ggsave(paste(fPath,'38462_Pskem_T_SubSeriesPlot.pdf',sep=""),
       device ="pdf",
       width  = 6,
       height = 2.5)
```

##### Quarterly T
```{r,message=FALSE,warning=FALSE}
pl <- data_long_tbl %>% filter(name=='T38462') %>% 
  summarise_by_time(.date_var = date, .by="quarter",value=mean(value)) %>% 
  tk_ts(frequency = 4) %>% 
  forecast::ggsubseriesplot(year.labels = FALSE) + 
              geom_smooth(method = "lm",color="red") +
              ggtitle('Decadal mean temperature (Station 38462, 1933 - 2015)') +
              xlab('month') +
              ylab('deg. C.')

pl

first_dec_T <- data_long_tbl %>% filter(name=='T38462') %>% summarise_by_time(.by = "quarter",meanT=mean(value)) %>% tk_ts(frequency = 4) %>% as.numeric() %>% head(40) %>% 
  pracma::Reshape(.,10,4) %>% colMeans()

last_dec_T <- data_long_tbl %>% filter(name=='T38462') %>% summarise_by_time(.by = "quarter",meanT=mean(value)) %>% tk_ts(frequency = 4) %>% as.numeric() %>% tail(40) %>% 
  pracma::Reshape(.,10,4) %>% colMeans()

delta_T <- last_dec_T - first_dec_T

delta_T_perc <- round(delta_T / first_dec_T * 100 ,digits = 0)
delta_T_perc

# fPath <-  '/Users/tobiassiegfried/Dropbox (hydrosolutions)/1_HSOL_PROJECTS/PROJECTS/SDC/DKU_WRM_COURSE_CA/Course Materials/Data/METEO/ChirchikRiverBasin/'
# ggsave(paste(fPath,'38462_Pskem_T_SubSeriesPlot.pdf',sep=""),
#        device ="pdf",
#        width  = 6,
#        height = 2.5)
```

#### 38462 - P
P: Pskem Meteorological Station (38462)
##### Monthly
```{r,message=FALSE,warning=FALSE}
pl <- data_long_tbl %>% filter(name=='P38462') %>% 
  summarise_by_time(.date_var = date, .by="month",value=sum(value)) %>% 
  tk_ts(frequency = 12) %>% 
  forecast::ggsubseriesplot(year.labels = FALSE) + 
              geom_smooth(method = "lm",color="red") +
              ggtitle('Mon. precipitation (Station 38462, 1933 - 2015)') +
              xlab('month') +
              ylab('mm')

fPath <-  '/Users/tobiassiegfried/Dropbox (hydrosolutions)/1_HSOL_PROJECTS/PROJECTS/SDC/DKU_WRM_COURSE_CA/Course Materials/Data/METEO/ChirchikRiverBasin/'
ggsave(paste(fPath,'38462_Pskem_P_SubSeriesPlot.pdf',sep=""),
       device ="pdf",
       width  = 6,
       height = 2.5)
```

##### Quarter
```{r,message=FALSE,warning=FALSE}
pl <- data_long_tbl %>% filter(name=='P38462') %>% 
  summarise_by_time(.date_var = date, .by="quarter",value=sum(value)) %>% 
  tk_ts(frequency = 4) %>% 
  forecast::ggsubseriesplot(year.labels = FALSE) + 
              geom_smooth(method = "lm",color="red") +
              ggtitle('Mon. precipitation (Station 38462, 1933 - 2015)') +
              xlab('month') +
              ylab('mm')

pl


data_long_tbl %>% filter(name=='P38462') %>% summarise_by_time(.by = "quarter",totP=sum(value)) %>% tk_ts(frequency = 4)

# fPath <-  '/Users/tobiassiegfried/Dropbox (hydrosolutions)/1_HSOL_PROJECTS/PROJECTS/SDC/DKU_WRM_COURSE_CA/Course Materials/Data/METEO/ChirchikRiverBasin/'
# ggsave(paste(fPath,'38462_Pskem_P_SubSeriesPlot.pdf',sep=""),
#        device ="pdf",
#        width  = 6,
#        height = 2.5)
```




<!-- ### Oygaing Met Station (38339) -->

<!-- ```{r,message=FALSE,warning=FALSE} -->
<!-- data %>% filter(code=='38339') %>% summarise_by_time(.date_var = date, .by="month",data=sum(data)) %>% -->
<!--     tk_ts(frequency = 12) %>% -->
<!--   forecast::ggsubseriesplot(year.labels = FALSE) + -->
<!--               geom_smooth(method = "lm",color="red") + -->
<!--               ggtitle('Mon. precipitation (Station 38339, 1962 - 2015)') + -->
<!--               xlab('month') + -->
<!--               ylab('mm') -->

<!-- ``` -->


### Chatkal River
#### 16279 - Q
##### Monthly
```{r,message=FALSE,warning=FALSE}
pl <- data_long_tbl %>% filter(name=='Q16279') %>% 
  summarise_by_time(.date_var = date, .by="month",value=mean(value)) %>% 
  tk_ts(frequency = 12) %>% 
  forecast::ggsubseriesplot(year.labels = FALSE) + 
              geom_smooth(method = "lm",color="red") +
              ggtitle('Mean mon. discharge (Station 16279, 1933 - 2015)') +
              xlab('month') +
              ylab('m^3/s')

fPath <-  '/Users/tobiassiegfried/Dropbox (hydrosolutions)/1_HSOL_PROJECTS/PROJECTS/SDC/DKU_WRM_COURSE_CA/Course Materials/Data/Q/ChirchikRiverBasin/'
ggsave(paste(fPath,'16279_Chatkal_Q_SubSeriesPlot.pdf',sep=""),
       device ="pdf",
       width  = 6,
       height = 2.5)

```
##### Quarter
```{r,message=FALSE,warning=FALSE}
pl <- data_long_tbl %>% filter(name=='Q16279') %>% 
  summarise_by_time(.date_var = date, .by="quarter",value=mean(value)) %>% 
  tk_ts(frequency = 4) %>% 
  forecast::ggsubseriesplot(year.labels = FALSE) + 
              geom_smooth(method = "lm",color="red") +
              ggtitle('Mean mon. discharge (Station 16279, 1933 - 2015)') +
              xlab('month') +
              ylab('mm')

pl

# fPath <-  '/Users/tobiassiegfried/Dropbox (hydrosolutions)/1_HSOL_PROJECTS/PROJECTS/SDC/DKU_WRM_COURSE_CA/Course Materials/Data/Q/ChirchikRiverBasin/'
# ggsave(paste(fPath,'16279_Chatkal_Q_SubSeriesPlot.pdf',sep=""),
#        device ="pdf",
#        width  = 6,
#        height = 2.5)

```

Get long term average Chatkal River discharge
```{r}
data_long_tbl %>% filter(name=='Q16279') %>% summarise(Average=mean(value,na.rm=TRUE))
```
#### 38471 - T
T: Chatkal Meteorological Station (38471)
##### Month
```{r,message=FALSE,warning=FALSE}
pl <- data_long_tbl %>% filter(name=='T38471') %>% 
  summarise_by_time(.date_var = date, .by="month",value=mean(value)) %>% 
  tk_ts(frequency = 12) %>% 
  forecast::ggsubseriesplot(year.labels = FALSE) + 
              geom_smooth(method = "lm",color="red") +
              ggtitle('Mon. mean temperature (Station 38471, 1933 - 2015)') +
              xlab('month') +
              ylab('deg. C.')

fPath <-  '/Users/tobiassiegfried/Dropbox (hydrosolutions)/1_HSOL_PROJECTS/PROJECTS/SDC/DKU_WRM_COURSE_CA/Course Materials/Data/METEO/ChirchikRiverBasin/'
ggsave(paste(fPath,'38471_Chatkal_T_SubSeriesPlot.pdf',sep=""),
       device ="pdf",
       width  = 6,
       height = 2.5)
```
##### Quarter
```{r,message=FALSE,warning=FALSE}
pl <- data_long_tbl %>% filter(name=='T38471') %>% 
  summarise_by_time(.date_var = date, .by="quarter",value=mean(value)) %>% 
  tk_ts(frequency = 4) %>% 
  forecast::ggsubseriesplot(year.labels = FALSE) + 
              geom_smooth(method = "lm",color="red") +
              ggtitle('Mon. mean temperature (Station 38471, 1933 - 2015)') +
              xlab('month') +
              ylab('deg. C.')

pl

data_long_tbl %>% filter(name=='T38471') %>% summarise_by_time(.by = "quarter",meanT=mean(value)) %>% tk_ts(frequency = 4)

# fPath <-  '/Users/tobiassiegfried/Dropbox (hydrosolutions)/1_HSOL_PROJECTS/PROJECTS/SDC/DKU_WRM_COURSE_CA/Course Materials/Data/METEO/ChirchikRiverBasin/'
# ggsave(paste(fPath,'38471_Chatkal_T_SubSeriesPlot.pdf',sep=""),
#        device ="pdf",
#        width  = 6,
#        height = 2.5)
```


#### 38471 - P
P: Chatkal Meteorological Station (38471)
##### Monthly
```{r,message=FALSE,warning=FALSE}
pl <- data_long_tbl %>% filter(name=='P38471') %>% 
  summarise_by_time(.date_var = date, .by="month",value=sum(value)) %>% 
  tk_ts(frequency = 12) %>% 
  forecast::ggsubseriesplot(year.labels = FALSE) + 
              geom_smooth(method = "lm",color="red") +
              ggtitle('Mon. precipitation (Station 38471, 1933 - 2015)') +
              xlab('month') +
              ylab('mm')

fPath <-  '/Users/tobiassiegfried/Dropbox (hydrosolutions)/1_HSOL_PROJECTS/PROJECTS/SDC/DKU_WRM_COURSE_CA/Course Materials/Data/METEO/ChirchikRiverBasin/'
ggsave(paste(fPath,'38471_Chatkal_P_SubSeriesPlot.pdf',sep=""),
       device ="pdf",
       width  = 6,
       height = 2.5)
```

##### Quarter
```{r,message=FALSE,warning=FALSE}
pl <- data_long_tbl %>% filter(name=='P38471') %>% 
  summarise_by_time(.date_var = date, .by="quarter",value=sum(value)) %>% 
  tk_ts(frequency = 4) %>% 
  forecast::ggsubseriesplot(year.labels = FALSE) + 
              geom_smooth(method = "lm",color="red") +
              ggtitle('Mon. precipitation (Station 38471, 1933 - 2015)') +
              xlab('month') +
              ylab('mm')

pl

data_long_tbl %>% filter(name=='P38471') %>% summarise_by_time(.by = "quarter",totP=sum(value)) %>% tk_ts(frequency = 4)

# fPath <-  '/Users/tobiassiegfried/Dropbox (hydrosolutions)/1_HSOL_PROJECTS/PROJECTS/SDC/DKU_WRM_COURSE_CA/Course Materials/Data/METEO/ChirchikRiverBasin/'
# ggsave(paste(fPath,'38471_Chatkal_P_SubSeriesPlot.pdf',sep=""),
#        device ="pdf",
#        width  = 6,
#        height = 2.5)
```

#### Ugam
##### 16300 - Q
```{r,message=FALSE,warning=FALSE}
data <- ChirchikRiverBasin
ugam <- data %>% filter(code=='16300')
pl <- ugam %>% 
  summarise_by_time(.date_var = date, .by="month",value=mean(data)) %>% 
  tk_ts(frequency = 12) %>% 
  forecast::ggsubseriesplot(year.labels = FALSE) + 
              geom_smooth(method = "lm",color="red") +
              ggtitle('Monthly mean discharge (Station 16300, 1932 - 2011)') +
              xlab('month') +
              ylab('deg. C.')

ugam %>% plot_time_series(date,data,.smooth = FALSE)

pl
```

#### Nauvalisoy
##### 16298 - Q
```{r,message=FALSE,warning=FALSE}
data <- ChirchikRiverBasin
nauvalisoy <- data %>% filter(code=='16298')
pl <- nauvalisoy %>% 
  summarise_by_time(.date_var = date, .by="month",value=mean(data)) %>% 
  tk_ts(frequency = 12) %>% 
  forecast::ggsubseriesplot(year.labels = FALSE) + 
              geom_smooth(method = "lm",color="red") +
              ggtitle('Monthly mean discharge (Station 16298, 1963 - 2015)') +
              xlab('month') +
              ylab('deg. C.')

nauvalisoy %>% plot_time_series(date,data,.smooth = FALSE)


pl
```


## Syr Darya

### Kara Darya

#### 16938 - Q
##### Monthly
```{r,message=FALSE,warning=FALSE}
fPath <-  '/Users/tobiassiegfried/Dropbox (hydrosolutions)/1_HSOL_PROJECTS/PROJECTS/SDC/DKU_WRM_COURSE_CA/Course Materials/Data/Q/SyrDarya/'
fName <-  '16938_decadal.csv'
Q16938 <- loadTabularData(fPath,fName,'16938','KaraDarya','KaraDarya','SyrDarya','Q','mm') 

Q16938 %>% 
  summarise_by_time(.date_var = date, .by="month",data=mean(data)) %>% plot_time_series(date,data)

pl <- Q16938  %>% summarise_by_time(.date_var = date, .by="month",data=mean(data)) %>% 
  tk_ts(frequency = 12) %>% 
  forecast::ggsubseriesplot(year.labels = FALSE) + 
              geom_smooth(method = "lm",color="red") +
              ggtitle('Mean mon. discharge (Station 16938, 1926 - 2017)') +
              xlab('month') +
              ylab('m^3/s')

ggsave(paste(fPath,'16938_KaraDara_SubSeriesPlot.pdf',sep=""),
       device ="pdf",
       width  = 6,
       height = 2.5)
```

##### Quarter
```{r,message=FALSE,warning=FALSE}
fPath <-  '/Users/tobiassiegfried/Dropbox (hydrosolutions)/1_HSOL_PROJECTS/PROJECTS/SDC/DKU_WRM_COURSE_CA/Course Materials/Data/Q/SyrDarya/'
fName <-  '16938_decadal.csv'
Q16938 <- loadTabularData(fPath,fName,'16938','KaraDarya','KaraDarya','SyrDarya','Q','mm') 

Q16938 %>% 
  summarise_by_time(.date_var = date, .by="month",data=mean(data)) %>% plot_time_series(date,data)

pl <- Q16938  %>% summarise_by_time(.date_var = date, .by="quarter",data=mean(data)) %>% 
  tk_ts(frequency = 4) %>% 
  forecast::ggsubseriesplot(year.labels = FALSE) + 
              geom_smooth(method = "lm",color="red") +
              ggtitle('Mean mon. discharge (Station 16938, 1926 - 2017)') +
              xlab('month') +
              ylab('m^3/s')

pl

# ggsave(paste(fPath,'16938_KaraDara_SubSeriesPlot.pdf',sep=""),
#        device ="pdf",
#        width  = 6,
#        height = 2.5)
```


#### 38476 - P 
P: Ak Terek Gava 38476
##### Monthly
```{r,message=FALSE,warning=FALSE}
detach("package:riversCentralAsia", unload=TRUE)
library(riversCentralAsia)

fPath <-  '/Users/tobiassiegfried/Dropbox (hydrosolutions)/1_HSOL_PROJECTS/PROJECTS/SDC/DKU_WRM_COURSE_CA/Course Materials/Data/METEO/SyrDarya/'
fName <-  '38476_P_decadal.csv'
P38476 = loadTabularData(fPath,fName,'38476','Ak-TerekGava','KaraDarya','SyDarya','P','mm')

pl <- P38476 %>% summarise_by_time(.date_var = date, .by="month",data=sum(data)) %>% 
  tk_ts(frequency = 12) %>% 
  forecast::ggsubseriesplot(year.labels = FALSE) + 
              geom_smooth(method = "lm",color="red") +
              ggtitle('Total Monthly Precipitation (1940 - 2016), Ak-Terek Gava, KaraDarya') +
              xlab('month') +
              ylab('mm')

ggsave(paste(fPath,'38476_AkTerekGava_P_SubSeriesPlot.pdf',sep=""),
       device="pdf",
       width = 6,
       height= 2.5)
```

##### Quarterly
```{r,message=FALSE,warning=FALSE}
detach("package:riversCentralAsia", unload=TRUE)
library(riversCentralAsia)

fPath <-  '/Users/tobiassiegfried/Dropbox (hydrosolutions)/1_HSOL_PROJECTS/PROJECTS/SDC/DKU_WRM_COURSE_CA/Course Materials/Data/METEO/SyrDarya/'
fName <-  '38476_P_decadal.csv'
P38476 = loadTabularData(fPath,fName,'38476','Ak-TerekGava','KaraDarya','SyDarya','P','mm')

P38476 %>% summarise_by_time(.by = "quarter",sumP=sum(data)) %>% tk_ts(frequency = 4)


pl <- P38476 %>% summarise_by_time(.date_var = date, .by="quarter",data=sum(data)) %>% 
  tk_ts(frequency = 4) %>% 
  forecast::ggsubseriesplot(year.labels = FALSE) + 
              geom_smooth(method = "lm",color="red") +
              ggtitle('Total Monthly Precipitation (1940 - 2016), Ak-Terek Gava, KaraDarya') +
              xlab('month') +
              ylab('mm')

pl

# ggsave(paste(fPath,'38476_AkTerekGava_P_SubSeriesPlot.pdf',sep=""),
#        device="pdf",
#        width = 6,
#        height= 2.5)
```


#### 38476 - T
##### Monthly
```{r,message=FALSE,warning=FALSE}
detach("package:riversCentralAsia", unload=TRUE)
library(riversCentralAsia)

fPath <-  '/Users/tobiassiegfried/Dropbox (hydrosolutions)/1_HSOL_PROJECTS/PROJECTS/SDC/DKU_WRM_COURSE_CA/Course Materials/Data/METEO/SyrDarya/'
fName <-  '38476_T_decadal.csv'
T38476 = loadTabularData(fPath,fName,'38476','Ak-TerekGava','KaraDarya','SyDarya','T','degC')

pl <- T38476 %>% summarise_by_time(.date_var = date, .by="month",data=mean(data)) %>% 
  tk_ts(frequency = 12) %>% 
  forecast::ggsubseriesplot(year.labels = FALSE) + 
              geom_smooth(method = "lm",color="red") +
              ggtitle('Mean Monthly Temperature (1940 - 2016), Ak-Terek Gava, KaraDarya') +
              xlab('month') +
              ylab('deg. C.')

ggsave(paste(fPath,'38476_AkTerekGava_T_SubSeriesPlot.pdf',sep=""),
       device="pdf",
       width = 6,
       height = 2.5)
```

##### Quarterly
```{r,message=FALSE,warning=FALSE}
detach("package:riversCentralAsia", unload=TRUE)
library(riversCentralAsia)

fPath <-  '/Users/tobiassiegfried/Dropbox (hydrosolutions)/1_HSOL_PROJECTS/PROJECTS/SDC/DKU_WRM_COURSE_CA/Course Materials/Data/METEO/SyrDarya/'
fName <-  '38476_T_decadal.csv'
T38476 = loadTabularData(fPath,fName,'38476','Ak-TerekGava','KaraDarya','SyDarya','T','degC')

pl <- T38476 %>% summarise_by_time(.date_var = date, .by="quarter",data=mean(data)) %>% 
  tk_ts(frequency = 4) %>% 
  forecast::ggsubseriesplot(year.labels = FALSE) + 
              geom_smooth(method = "lm",color="red") +
              ggtitle('Mean Monthly Temperature (1940 - 2016), Ak-Terek Gava, KaraDarya') +
              xlab('month') +
              ylab('deg. C.')

pl

T38476 %>% summarise_by_time(.by = "quarter",meanT=mean(data)) %>% tk_ts(frequency = 4)

# ggsave(paste(fPath,'38476_AkTerekGava_T_SubSeriesPlot.pdf',sep=""),
#        device="pdf",
#        width = 6,
#        height = 2.5)
```


### Sokh River

#### 16198 - Q
##### Monthly
```{r,message=FALSE,warning=FALSE}
fPath <-  '/Users/tobiassiegfried/Dropbox (hydrosolutions)/1_HSOL_PROJECTS/PROJECTS/SDC/DKU_WRM_COURSE_CA/Course Materials/Data/Q/SyrDarya/'
fName <-  '16198_decadal.csv'
Q16198 <- loadTabularData(fPath,fName,'16198','Sarykanda','Sokh','SyrDarya','Q','mm') 

pl <- Q16198  %>% summarise_by_time(.date_var = date, .by="month",data=mean(data)) %>% 
  tk_ts(frequency = 12) %>% 
  forecast::ggsubseriesplot(year.labels = FALSE) + 
              geom_smooth(method = "lm",color="red") +
              ggtitle('Mean mon. discharge (Station 16198, 1926 - 2017)') +
              xlab('month') +
              ylab('m^3/s')

ggsave(paste(fPath,'16198_Sokh_Q_SubSeriesPlot.pdf',sep=""),
       device ="pdf",
       width  = 6,
       height = 2.5)

```

##### Quarter
```{r,message=FALSE,warning=FALSE}
fPath <-  '/Users/tobiassiegfried/Dropbox (hydrosolutions)/1_HSOL_PROJECTS/PROJECTS/SDC/DKU_WRM_COURSE_CA/Course Materials/Data/Q/SyrDarya/'
fName <-  '16198_decadal.csv'
Q16198 <- loadTabularData(fPath,fName,'16198','Sarykanda','Sokh','SyrDarya','Q','mm') 

pl <- Q16198  %>% summarise_by_time(.date_var = date, .by="quarter",data=mean(data)) %>% 
  tk_ts(frequency = 4) %>% 
  forecast::ggsubseriesplot(year.labels = FALSE) + 
              geom_smooth(method = "lm",color="red") +
              ggtitle('Mean mon. discharge (Station 16198, 1926 - 2017)') +
              xlab('month') +
              ylab('m^3/s')

pl
# 
# ggsave(paste(fPath,'16198_Sokh_Q_SubSeriesPlot.pdf',sep=""),
#        device ="pdf",
#        width  = 6,
#        height = 2.5)

```


Get long term average Sokh River discharge
```{r}
Q16198 %>% summarise(Average=mean(data,na.rm=TRUE))
```
#### Abramov Met Station - T
##### Monthly
For Amu Darya, we have the complete climate record of the old Abramov Glacier Station that was destroyed in 1999 by Islamist Extremists.
```{r,message=FALSE,warning=FALSE}
fPath <- '/Users/tobiassiegfried/Dropbox (hydrosolutions)/1_HSOL_PROJECTS/PROJECTS/SDC/DKU_WRM_COURSE_CA/Course Materials/Data/METEO/AmuDarya/AmbramovStationData.csv'
abramovMet <- read_csv(fPath,col_names = TRUE) %>% tk_tbl()

abramovMet_Tav_mon <- abramovMet %>% 
  summarise_by_time(.date_var = Date,.by="month",Avg_Air_temp=mean(Avg_Air_temp)) %>% 
  filter(Date > as.Date('1967-12-01'))
abramovMet_Tav_mon %>% plot_time_series(Date,Avg_Air_temp,.smooth = FALSE,.interactive = FALSE)
pl <- abramovMet_Tav_mon %>%   tk_ts(frequency = 12) %>% na.omit %>% 
  forecast::ggsubseriesplot(year.labels = TRUE) + 
              geom_smooth(method = "lm",color="red") +
              ggtitle('Mon. mean temperature (Station Abramov, 1967 - 1999)') +
              xlab('month') +
              ylab('deg. C.')

ggsave(paste(fPath,'Abramov_T_SubSeriesPlot.pdf',sep=""),
       device ="pdf",
       width  = 6,
       height = 2.5)
```

##### Quarterly
For Amu Darya, we have the complete climate record of the old Abramov Glacier Station that was destroyed in 1999 by Islamist Extremists.
```{r,message=FALSE,warning=FALSE}
fPath <- '/Users/tobiassiegfried/Dropbox (hydrosolutions)/1_HSOL_PROJECTS/PROJECTS/SDC/DKU_WRM_COURSE_CA/Course Materials/Data/METEO/AmuDarya/AmbramovStationData.csv'
abramovMet <- read_csv(fPath,col_names = TRUE) %>% tk_tbl()

abramovMet_Tav_mon <- abramovMet %>% 
  summarise_by_time(.date_var = Date,.by="quarter",Avg_Air_temp=mean(Avg_Air_temp)) %>% 
  filter(Date > as.Date('1967-12-01'))
# abramovMet_Tav_mon %>% plot_time_series(Date,Avg_Air_temp,.smooth = FALSE,.interactive = FALSE)
pl <- abramovMet_Tav_mon %>%   tk_ts(frequency = 4) %>% na.omit %>% 
  forecast::ggsubseriesplot(year.labels = TRUE) + 
              geom_smooth(method = "lm",color="red") +
              ggtitle('Mon. mean temperature (Station Abramov, 1967 - 1999)') +
              xlab('month') +
              ylab('deg. C.')

pl

abramovMet %>% summarise_by_time(.date_var = Date,.by="quarter",Avg_Air_temp=mean(Avg_Air_temp)) %>%  filter(Date > as.Date('1967-12-01')) %>%   tk_ts(frequency = 4)

# ggsave(paste(fPath,'Abramov_T_SubSeriesPlot.pdf',sep=""),
#        device ="pdf",
#        width  = 6,
#        height = 2.5)
```

#### Abramov Met Station - P
##### Monthly
```{r,message=FALSE,warning=FALSE}
fPath <- '/Users/tobiassiegfried/Dropbox (hydrosolutions)/1_HSOL_PROJECTS/PROJECTS/SDC/DKU_WRM_COURSE_CA/Course Materials/Data/METEO/AmuDarya/AmbramovStationData.csv'
abramovMet <- read_csv(fPath,col_names = TRUE) %>% tk_tbl()

abramovMet_Psum_mon <- abramovMet %>% 
  summarise_by_time(.date_var = Date,.by="month",Precipitation=sum(Precipitation,na.rm = TRUE)) %>% 
  filter(Date > as.Date('1967-12-01'))
abramovMet_Psum_mon %>% plot_time_series(Date,Precipitation,.smooth = FALSE,.interactive = FALSE)
pl <- abramovMet_Psum_mon %>%   tk_ts(frequency = 12) %>% na.omit %>% 
  forecast::ggsubseriesplot(year.labels = TRUE) + 
              geom_smooth(method = "lm",color="red") +
              ggtitle('Mon. precipitation (Station Abramov, 1967 - 1999)') +
              xlab('month') +
              ylab('mm')

ggsave(paste(fPath,'Abramov_P_SubSeriesPlot.pdf',sep=""),
       device ="pdf",
       width  = 6,
       height = 2.5)
```

##### Quarterly
```{r,message=FALSE,warning=FALSE}
fPath <- '/Users/tobiassiegfried/Dropbox (hydrosolutions)/1_HSOL_PROJECTS/PROJECTS/SDC/DKU_WRM_COURSE_CA/Course Materials/Data/METEO/AmuDarya/AmbramovStationData.csv'
abramovMet <- read_csv(fPath,col_names = TRUE) %>% tk_tbl()

abramovMet_Psum_mon <- abramovMet %>% 
  summarise_by_time(.date_var = Date,.by="quarter",Precipitation=sum(Precipitation,na.rm = TRUE)) %>% 
  filter(Date > as.Date('1967-12-01'))
# abramovMet_Psum_mon %>% plot_time_series(Date,Precipitation,.smooth = FALSE,.interactive = FALSE)
pl <- abramovMet_Psum_mon %>%   tk_ts(frequency = 4) %>% na.omit %>% 
  forecast::ggsubseriesplot(year.labels = TRUE) + 
              geom_smooth(method = "lm",color="red") +
              ggtitle('Mon. precipitation (Station Abramov, 1967 - 1999)') +
              xlab('month') +
              ylab('mm')

pl

abramovMet %>% 
  summarise_by_time(.date_var = Date,.by="quarter",Precipitation=sum(Precipitation,na.rm = TRUE)) %>% 
  filter(Date > as.Date('1967-12-01')) %>% tk_ts(frequency = 4)

# ggsave(paste(fPath,'Abramov_P_SubSeriesPlot.pdf',sep=""),
#        device ="pdf",
#        width  = 6,
#        height = 2.5)
```


## Chu River

### Ala Archa

#### 15194 - Q
##### Monthly
```{r,message=FALSE,warning=FALSE}
fPath <-  '/Users/tobiassiegfried/Dropbox (hydrosolutions)/1_HSOL_PROJECTS/PROJECTS/SDC/DKU_WRM_COURSE_CA/Course Materials/Data/Q/ChuRiver/'
fName <-  'AlaArcha.csv'
AlaArcha_Q = loadTabularData(fPath,fName,'15194','AlaArcha','AlaArcha','ChuRiver','Q','mm')

pl <- AlaArcha_Q  %>% summarise_by_time(.date_var = date, .by="month",data=mean(data)) %>% 
  tk_ts(frequency = 12) %>% 
  forecast::ggsubseriesplot(year.labels = FALSE) + 
              geom_smooth(method = "lm",color="red") +
              ggtitle('Mean mon. discharge (Station 15194, 1936 - 2014)') +
              xlab('month') +
              ylab('m^3/s')

ggsave(paste(fPath,'15194_AlaArcha_Q_SubSeriesPlot.pdf',sep=""),
       device ="pdf",
       width  = 6,
       height = 2.5)

```
##### Quarter
```{r,message=FALSE,warning=FALSE}
fPath <-  '/Users/tobiassiegfried/Dropbox (hydrosolutions)/1_HSOL_PROJECTS/PROJECTS/SDC/DKU_WRM_COURSE_CA/Course Materials/Data/Q/ChuRiver/'
fName <-  'AlaArcha.csv'
AlaArcha_Q = loadTabularData(fPath,fName,'15194','AlaArcha','AlaArcha','ChuRiver','Q','mm')

pl <- AlaArcha_Q  %>% summarise_by_time(.date_var = date, .by="quarter",data=mean(data)) %>% 
  tk_ts(frequency = 4) %>% 
  forecast::ggsubseriesplot(year.labels = FALSE) + 
              geom_smooth(method = "lm",color="red") +
              ggtitle('Mean mon. discharge (Station 15194, 1936 - 2014)') +
              xlab('month') +
              ylab('m^3/s')

pl

# 
# ggsave(paste(fPath,'15194_AlaArcha_Q_SubSeriesPlot.pdf',sep=""),
#        device ="pdf",
#        width  = 6,
#        height = 2.5)

```


Get long term average Ala Archa River discharge
```{r}
AlaArcha_Q %>% summarise(Average=mean(data,na.rm=TRUE))
```

#### Baytik - P
##### Monthly P
```{r,message=FALSE}
detach("package:riversCentralAsia", unload=TRUE)
library(riversCentralAsia)

fPath <-  '/Users/tobiassiegfried/Dropbox (hydrosolutions)/1_HSOL_PROJECTS/PROJECTS/SDC/DKU_WRM_COURSE_CA/Course Materials/Data/METEO/ChuRiver/'
fName <-  'Baytik-P_clean.csv'
baytik_P = loadTabularData(fPath,fName,'11111','Baytik','Ala-Archa','ChuRiver','P','degC')

pl <- baytik_P %>% na.omit() %>% summarise_by_time(.date_var = date, .by="month",data=sum(data)) %>% 
  tk_ts(frequency = 12) %>% 
  forecast::ggsubseriesplot(year.labels = FALSE) + 
              geom_smooth(method = "lm",color="red") +
              ggtitle('Mon. precipitation (Station Baytik, 1926 - 2019)') +
              xlab('month') +
              ylab('mm')

ggsave(paste(fPath,'Baytik_AlaArcha_P_SubSeriesPlot.pdf',sep=""),
       device ="pdf",
       width  = 6,
       height = 2.5)

```

##### Quarterly P
```{r,message=FALSE}
detach("package:riversCentralAsia", unload=TRUE)
library(riversCentralAsia)

fPath <-  '/Users/tobiassiegfried/Dropbox (hydrosolutions)/1_HSOL_PROJECTS/PROJECTS/SDC/DKU_WRM_COURSE_CA/Course Materials/Data/METEO/ChuRiver/'
fName <-  'Baytik-P_clean.csv'
baytik_P = loadTabularData(fPath,fName,'11111','Baytik','Ala-Archa','ChuRiver','P','degC')

pl <- baytik_P %>% na.omit() %>% summarise_by_time(.date_var = date, .by="quarter",data=sum(data)) %>% 
  tk_ts(frequency = 4) %>% 
  forecast::ggsubseriesplot(year.labels = FALSE) + 
              geom_smooth(method = "lm",color="red") +
              ggtitle('Mon. precipitation (Station Baytik, 1926 - 2019)') +
              xlab('month') +
              ylab('mm')

pl

baytik_P %>% summarise_by_time(.by = "quarter",totP=sum(data)) %>% tk_ts(frequency = 4)

# ggsave(paste(fPath,'Baytik_AlaArcha_P_SubSeriesPlot.pdf',sep=""),
#        device ="pdf",
#        width  = 6,
#        height = 2.5)

```

#### Baytik - T
##### Monthly T
```{r,message=FALSE,warning=FALSE}
detach("package:riversCentralAsia", unload=TRUE)
library(riversCentralAsia)

fPath <-  '/Users/tobiassiegfried/Dropbox (hydrosolutions)/1_HSOL_PROJECTS/PROJECTS/SDC/DKU_WRM_COURSE_CA/Course Materials/Data/METEO/ChuRiver/'
fName <-  'Baytik-T_clean.csv'
baytik_T = loadTabularData(fPath,fName,'11111','Baytik','Ala-Archa','ChuRiver','T','degC')

pl <- baytik_T %>% na.omit() %>% summarise_by_time(.date_var = date, .by="month",data=mean(data)) %>% 
  tk_ts(frequency = 12) %>% 
  forecast::ggsubseriesplot(year.labels = FALSE) + 
              geom_smooth(method = "lm",color="red") +
              ggtitle('Mon. mean temperature (Station Baytik, 1926 - 2019)') +
              xlab('month') +
              ylab('deg. C.')

ggsave(paste(fPath,'Baytik_AlaArcha_T_SubSeriesPlot.pdf',sep=""),
       device ="pdf",
       width  = 6,
       height = 2.5)
```

##### Quarterly T
```{r,message=FALSE,warning=FALSE}
detach("package:riversCentralAsia", unload=TRUE)
library(riversCentralAsia)

fPath <-  '/Users/tobiassiegfried/Dropbox (hydrosolutions)/1_HSOL_PROJECTS/PROJECTS/SDC/DKU_WRM_COURSE_CA/Course Materials/Data/METEO/ChuRiver/'
fName <-  'Baytik-T_clean.csv'
baytik_T = loadTabularData(fPath,fName,'11111','Baytik','Ala-Archa','ChuRiver','T','degC')

pl <- baytik_T %>% na.omit() %>% summarise_by_time(.date_var = date, .by="quarter",data=mean(data)) %>% 
  tk_ts(frequency = 4) %>% 
  forecast::ggsubseriesplot(year.labels = FALSE) + 
              geom_smooth(method = "lm",color="red") +
              ggtitle('Mon. mean temperature (Station Baytik, 1926 - 2019)') +
              xlab('month') +
              ylab('deg. C.')

pl

baytik_T %>% summarise_by_time(.by = "quarter",meanT=mean(data)) %>% tk_ts(frequency = 4)


# ggsave(paste(fPath,'Baytik_AlaArcha_T_SubSeriesPlot.pdf',sep=""),
#        device ="pdf",
#        width  = 6,
#        height = 2.5)
```

#### Alamedin - Q
##### Monthly
```{r,message=FALSE,warning=FALSE}
fPath <-  '/Users/tobiassiegfried/Dropbox (hydrosolutions)/1_HSOL_PROJECTS/PROJECTS/SDC/DKU_WRM_COURSE_CA/Course Materials/Data/Q/ChuRiver/'
fName <-  'Alamedin.csv'
Alamedin_Q = loadTabularData(fPath,fName,'15189','Alamedin','Alamedin','ChuRiver','Q','mm')

pl <- Alamedin_Q  %>% summarise_by_time(.date_var = date, .by="month",data=mean(data)) %>% 
  tk_ts(frequency = 12) %>% 
  forecast::ggsubseriesplot(year.labels = FALSE) + 
              geom_smooth(method = "lm",color="red") +
              ggtitle('Mean mon. discharge (Station 15189 (Alamedin), 1911 - 2014)') +
              xlab('month') +
              ylab('m^3/s')

pl


# ggsave(paste(fPath,'15194_AlaArcha_Q_SubSeriesPlot.pdf',sep=""),
#        device ="pdf",
#        width  = 6,
#        height = 2.5)

```

#### KaraBalta - Q
##### Monthly
```{r,message=FALSE,warning=FALSE}
fPath <-  '/Users/tobiassiegfried/Dropbox (hydrosolutions)/1_HSOL_PROJECTS/PROJECTS/SDC/DKU_WRM_COURSE_CA/Course Materials/Data/Q/ChuRiver/'
fName <-  'KaraBalta.csv'
KaraBalta_Q = loadTabularData(fPath,fName,'15215','KaraBalta','KaraBalta','ChuRiver','Q','mm')

pl <- KaraBalta_Q  %>% summarise_by_time(.date_var = date, .by="month",data=mean(data)) %>% 
  tk_ts(frequency = 12) %>% 
  forecast::ggsubseriesplot(year.labels = FALSE) + 
              geom_smooth(method = "lm",color="red") +
              ggtitle('Mean mon. discharge (Station 15215 (KaraBalta), 1925 - 2014)') +
              xlab('month') +
              ylab('m^3/s')

pl


# ggsave(paste(fPath,'15194_AlaArcha_Q_SubSeriesPlot.pdf',sep=""),
#        device ="pdf",
#        width  = 6,
#        height = 2.5)

```

#### Kegety - Q
##### Monthly
```{r,message=FALSE,warning=FALSE}
fPath <-  '/Users/tobiassiegfried/Dropbox (hydrosolutions)/1_HSOL_PROJECTS/PROJECTS/SDC/DKU_WRM_COURSE_CA/Course Materials/Data/Q/ChuRiver/'
fName <-  'Kegety.csv'
Kegety_Q = loadTabularData(fPath,fName,'15171','Kegety','Kegety','ChuRiver','Q','mm')

pl <- Kegety_Q  %>% summarise_by_time(.date_var = date, .by="month",data=mean(data)) %>% 
  tk_ts(frequency = 12) %>% 
  forecast::ggsubseriesplot(year.labels = FALSE) + 
              geom_smooth(method = "lm",color="red") +
              ggtitle('Mean mon. discharge (Station 15171 (Kegety), 1925 - 2014)') +
              xlab('month') +
              ylab('m^3/s')

pl


# ggsave(paste(fPath,'15194_AlaArcha_Q_SubSeriesPlot.pdf',sep=""),
#        device ="pdf",
#        width  = 6,
#        height = 2.5)

```

#### Chon Kemin - Q
##### Monthly
```{r,message=FALSE,warning=FALSE}
fPath <-  '/Users/tobiassiegfried/Dropbox (hydrosolutions)/1_HSOL_PROJECTS/PROJECTS/SDC/DKU_WRM_COURSE_CA/Course Materials/Data/Q/ChuRiver/'
fName <-  'ChonKemin_2005.csv'
ChonKemin_Q = loadTabularData(fPath,fName,'15149','ChonKemin','ChonKemin','ChuRiver','Q','mm')

pl <- ChonKemin_Q  %>% summarise_by_time(.date_var = date, .by="month",data=mean(data)) %>% 
  tk_ts(frequency = 12) %>% 
  forecast::ggsubseriesplot(year.labels = FALSE) + 
              geom_smooth(method = "lm",color="red") +
              ggtitle('Mean mon. discharge (Station 15149 (ChonKemin), 1936 - 2014)') +
              xlab('month') +
              ylab('m^3/s')

pl


# ggsave(paste(fPath,'15194_AlaArcha_Q_SubSeriesPlot.pdf',sep=""),
#        device ="pdf",
#        width  = 6,
#        height = 2.5)

```


## Amu Darya

### Gund River

#### 17050 - Q
##### Monthly
```{r,message=FALSE,warning=FALSE}
fPath <-  '/Users/tobiassiegfried/Dropbox (hydrosolutions)/1_HSOL_PROJECTS/PROJECTS/SDC/DKU_WRM_COURSE_CA/Course Materials/Data/Q/AmuDarya/'
fName <-  '17050_Gund_Khorog_mon.csv'
Q17050 <- loadTabularData(fPath,fName,'17050','Khorog','Gunt','AmuDarya','Q','m^3/s')

pl <- Q17050%>% summarise_by_time(.date_var = date, .by="month",data=mean(data)) %>% 
  tk_ts(frequency = 12) %>% 
  forecast::ggsubseriesplot(year.labels = FALSE) + 
              geom_smooth(method = "lm",color="red") +
              ggtitle('Mean mon. discharge (Station 17050, 1940 - 2016)') +
              xlab('month') +
              ylab('m^3/s')

ggsave(paste(fPath,'17050_Gund_P_SubSeriesPlot.pdf',sep=""),
       device ="pdf",
       width  = 6,
       height = 2.5)

```

##### Quarter
```{r,message=FALSE,warning=FALSE}
fPath <-  '/Users/tobiassiegfried/Dropbox (hydrosolutions)/1_HSOL_PROJECTS/PROJECTS/SDC/DKU_WRM_COURSE_CA/Course Materials/Data/Q/AmuDarya/'
fName <-  '17050_Gund_Khorog_mon.csv'
Q17050 <- loadTabularData(fPath,fName,'17050','Khorog','Gunt','AmuDarya','Q','m^3/s')

pl <- Q17050%>% summarise_by_time(.date_var = date, .by="quarter",data=mean(data)) %>% 
  tk_ts(frequency = 4) %>% 
  forecast::ggsubseriesplot(year.labels = FALSE) + 
              geom_smooth(method = "lm",color="red") +
              ggtitle('Mean mon. discharge (Station 17050, 1940 - 2016)') +
              xlab('month') +
              ylab('m^3/s')

pl

# ggsave(paste(fPath,'17050_Gund_P_SubSeriesPlot.pdf',sep=""),
#        device ="pdf",
#        width  = 6,
#        height = 2.5)

```


#### 38954 - P
##### Monthly
```{r}
fPath <-  '/Users/tobiassiegfried/Dropbox (hydrosolutions)/1_HSOL_PROJECTS/PROJECTS/SDC/DKU_WRM_COURSE_CA/Course Materials/Data/METEO/AmuDarya/'
fName <-  '38954_Khorog_P.csv'
P38954 <- loadTabularData(fPath,fName,'38954','Khorog','Gunt','AmuDarya','P','mm')

pl <- P38954 %>% summarise_by_time(.date_var = date, .by="month",data=sum(data)) %>% 
  tk_ts(frequency = 12) %>% 
  forecast::ggsubseriesplot(year.labels = FALSE) + 
              geom_smooth(method = "lm",color="red") +
              ggtitle('Mean mon. precipitation (Station 38954, 1900 - 2018)') +
              xlab('month') +
              ylab('m^3/s')

ggsave(paste(fPath,'38954_Khorog_P_SubSeriesPlot.pdf',sep=""),
       device ="pdf",
       width  = 6,
       height = 2.5)
```

##### Quarter
```{r}
fPath <-  '/Users/tobiassiegfried/Dropbox (hydrosolutions)/1_HSOL_PROJECTS/PROJECTS/SDC/DKU_WRM_COURSE_CA/Course Materials/Data/METEO/AmuDarya/'
fName <-  '38954_Khorog_P.csv'
P38954 <- loadTabularData(fPath,fName,'38954','Khorog','Gunt','AmuDarya','P','mm')

pl <- P38954 %>% summarise_by_time(.date_var = date, .by="quarter",data=sum(data)) %>% 
  tk_ts(frequency = 4) %>% 
  forecast::ggsubseriesplot(year.labels = FALSE) + 
              geom_smooth(method = "lm",color="red") +
              ggtitle('Mean mon. precipitation (Station 38954, 1900 - 2018)') +
              xlab('month') +
              ylab('m^3/s')

pl

P38954 %>% summarise_by_time(.date_var = date, .by="quarter",data=sum(data)) %>% 
  tk_ts(frequency = 4)

# ggsave(paste(fPath,'38954_Khorog_P_SubSeriesPlot.pdf',sep=""),
#        device ="pdf",
#        width  = 6,
#        height = 2.5)
```

#### 38954 - T
##### Monthly
```{r}
fPath <-  '/Users/tobiassiegfried/Dropbox (hydrosolutions)/1_HSOL_PROJECTS/PROJECTS/SDC/DKU_WRM_COURSE_CA/Course Materials/Data/METEO/AmuDarya/'
fName <-  '38954_Khorog_T.csv'

#dataMat <- read_csv(paste(fPath,fName,sep=""), col_names = FALSE, col_types = cols())

T38954 <- loadTabularData(fPath,fName,'38954','Khorog','Gunt','AmuDarya','T','mm')

pl <- T38954 %>% summarise_by_time(.date_var = date, .by="month",data=mean(data)) %>% 
  tk_ts(frequency = 12) %>% 
  forecast::ggsubseriesplot(year.labels = FALSE) + 
              geom_smooth(method = "lm",color="red") +
              ggtitle('Mon. mean temperature (Station 38954, 1899 - 2018)') +
              xlab('month') +
              ylab('m^3/s')

ggsave(paste(fPath,'38954_Khorog_T_SubSeriesPlot.pdf',sep=""),
       device ="pdf",
       width  = 6,
       height = 2.5)
```

##### Quarterly
```{r}
fPath <-  '/Users/tobiassiegfried/Dropbox (hydrosolutions)/1_HSOL_PROJECTS/PROJECTS/SDC/DKU_WRM_COURSE_CA/Course Materials/Data/METEO/AmuDarya/'
fName <-  '38954_Khorog_T.csv'

#dataMat <- read_csv(paste(fPath,fName,sep=""), col_names = FALSE, col_types = cols())

T38954 <- loadTabularData(fPath,fName,'38954','Khorog','Gunt','AmuDarya','T','mm')

pl <- T38954 %>% summarise_by_time(.date_var = date, .by="quarter",data=mean(data)) %>% 
  tk_ts(frequency = 4) %>% 
  forecast::ggsubseriesplot(year.labels = FALSE) + 
              geom_smooth(method = "lm",color="red") +
              ggtitle('Quarterly mean temperature (Station 38954, 1899 - 2018)') +
              xlab('month') +
              ylab('deg. C.')

pl

T38954 %>% summarise_by_time(.date_var = date, .by="quarter",data=mean(data)) %>% 
  tk_ts(frequency = 4)

# ggsave(paste(fPath,'38954_Khorog_T_SubSeriesPlot.pdf',sep=""),
#        device ="pdf",
#        width  = 6,
#        height = 2.5)
```


